---
title: "Quick guide to obtain domestic and foreign land use emissions from changes in domestic TFP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{brief_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The function `get.responses()` does all the work. For instance, to get the  effects on cropland area and emissions of all the countries in the sample stemming from an increase of 3\% in TFP in Brazil, use:
```{r setup}
library(LUEandAGTFP)
brazil.effects <- get.responses(ctry="bra", tfp.shock =  1)
head(brazil.effects)
```
By default, `get.responses()` uses the point estimates of the cropland elasticities to changes in TFP use. The option `boot=TRUE` will estimate 10,000 replicates of each elasticity using bootstrapped parameter estimates as explained in the paper. The syntax is:
```{r}
brazil.effects.boot <- get.responses(ctry="bra", tfp.shock =  1, boot = TRUE)
head(brazil.effects.boot)
```
Notice the addition of the variable `boot` which keeps track of the bootstrap replicate. The user can deploy standard `R` code to estimate the uncertainty of any of the output variables (`perc.change.crop.area`, `change.crops_ha`, `emissions_GtCO2`) derived from parameter uncertainty. For instance:
```{r}
alpha <- 0.10
bra.emission.ci <- aggregate(emissions_GtCO2 ~ k,
                           brazil.effects.boot,
                           function(x) quantile(x,
                                                probs=c( alpha/2,0.5,(1-alpha/2)),
                                                names = TRUE))
head(bra.emission.ci)
```
gives the quantile-based 90\% confidence interval for the effect of TFP growth in Brazil on each country in the sample, including Brazil itself:
```{r}
bra.emission.ci[bra.emission.ci$k =="bra",]
```

Other manipulations are possible, for instance, the global effect of a 1\% TFP in Brazil's ag. on worldwide emissions (excluding Brazil) can be calculated as:
```{r}
## Sum over countries within each bootstrap replicate:
global.effects.brazil <- aggregate( emissions_GtCO2 ~ boot,
                                   brazil.effects.boot[brazil.effects.boot$k != 'bra'],
                                   FUN = 'sum')
## Get quantile=based 90\% confidence intervals:
quantile(global.effects.brazil$emissions_GtCO2,
         probs=c( alpha/2,0.5,(1-alpha/2)),
         names = TRUE)                                   
```
So, while the domestic effect of TFP growth in Brazil is
statistically indistinguishable from zero, the foreign effects are
expected to be negative.

 The elasticities as well as the emission factors and cropland area are included with the package (type `?elasticities` and `?emission_data.ave` for documentation}. Few examples:
 
Mexico's cropland is highly sensitive to technological progress in the US: 
```{r}
data(elasticities)
## Percentage change in Mexico's cropland when TFP in the US grows by 1% (bilateral elasticity):
with( elasticities, elasticities[i == "usa" & k == "mex",])
```
But the converse is not true:
```{r}
## Percentage change in US cropland when TFP in Mexico grows by 1% (bilateral elasticity)
with( elasticities, elasticities[i == "mex" & k == "usa",])
```
Argentina is an example of a country where technological progress is expected to increase domestic cropland (so-called Jevons' paradox):
```{r}
## Percentage change in Argentina's cropland following a domestic
## increase of TFP of 1% (domestic elasticity):
with( elasticities, elasticities[i == "arg" & k == "arg",])
```
India on the other hand is an example of a country where TFP growth is land saving (the so-called Borlaug's hypothesis), at least using the point estimates:
```{r}
with( elasticities, elasticities[i == "ind" & k == "ind",])
```
